Proof of uniform convergence for a cell-centered AP discretization of the hyperbolic heat equation on general meshes

### 2015-03-19 Corrected_version.tex
adding corrections

\begin{proposition} \label{heqtode}
There exists a constant $C>0$ such that 
\begin{equation*}
\|\V^{\eps}-\W^\eps\|_{L^\infty([0,T];L^2(\mathbb{R}) )} \leq
C \eps  \|\partial_{xxx}p_0\|_{L^2(\mathbb{R})}.
\end{equation*}
\end{proposition}
\begin{proof}
We define $u^{\eps}=-\frac{\varepsilon}\sigma \partial_x p^{\eps}$ and
introduce $R^{\eps}$ such that the solution of the diffusion equation satisfies
\begin{equation}\label{modifdiff}
\left \{
\begin{array}{lll}
\dt p^{\eps} +\frac{1}{\eps}\partial_{x}u^{\eps}=0,\\
\partial_t u^{\eps} + \frac{1}{\eps}\partial_x p^{\eps} +\frac{\sigma}{\eps^2}u^{\eps} = R^{\eps}
\end{array}
\right.
\end{equation}
where
 $R^{\eps}=\dt u^{\eps}=-\frac{\varepsilon}\sigma \partial_{tx} p^{\eps}=
-\frac{\varepsilon}{\sigma^2} \partial_{xxx}p^{\eps}$. 
Note that 
$\|R^{\eps}(t)\|_{L^2(\mathbb{R})}\leq \|R^{\eps}(0)\|_{L^2(\mathbb{R})}$. 
Denoting $e^{\eps}=p-p^{\eps}$, $f^{\eps}=u-u^{\eps}$, we make the difference between the
systems \eqref{chaleur22} et \eqref{modifdiff}
\begin{equation}\label{diffmodif}
\left \{
\begin{array}{lll}
\dt e^{\eps} +\frac{1}{\eps}\partial_{x}f^{\eps}=0,\\
\partial_t f^{\eps} +\frac{1}{\eps}\partial_x e^{\eps} +\frac{\sigma}{\eps^2}f^{\eps} = R^{\eps}.
\end{array}
\right.
\end{equation}
Since data are well-prepared, one has $e^{\eps}(0)=f^{\eps}(0)=0$.
Consider
$\|\V^{\eps}-\W^\eps\|_{L^2(\mathbb{R})}^2=\|e^{\eps}\|_{L^2(\mathbb{R})}
^2+\|f^{\eps}\|_{L^2(\mathbb{R})}^2$.
Adding the first equation of
\eqref{diffmodif} multiplied by $e^{\eps}$ and the second multiplied by $f^{\eps}$ and integrating
on $\mathbb{R}$,
 we find out that:
$\frac12 \frac{d}{dt}\|\V^{\eps}-\W^\eps\|_{L^2(\mathbb{R})}^2 \leq
\int_{\mathbb{R}}R^{\eps}f^{\eps} dx \leq  \|R^{\eps}\|_{L^2(\mathbb{R})}
\|\V^{\eps}-\W^\eps\|_{L^2(\mathbb{R})}
.$
  The proof is ended by integration between 0 and T.

@@

\| \mathbf u^\eps  \|^2_{ L^2( [0,t] : H^1(\mathbb R) }.

Since $\| \mathbf u^\eps  \|^2_{ L^2( [0,t] : H^1(\mathbb R) }\leq t \| \V^{\eps}(0) \|_{H^1(\mathbb{R})}^2$ and also 
$
\frac{\sigma}
{\eps^2}\| \mathbf u^\eps  \|^2_{ L^2( [0,t] : H^1(\mathbb R) }\leq  \| \V^{\eps}(0) \|_{H^1(\mathbb{R})}^2$
by (\ref{bee1} ) and (\ref{bee2} ), one gets that
\begin{equation} \label{eq:sm1}
\left( \frac{1}{\eps}+\frac{\sigma}{2\eps^2}\right)
\int_0^t 
\sum_{j\in \mathbb{Z}} 
\left ( 
\delta_{j}^+(u^{\eps})^2+\delta_{j}^-(u^{\eps})^2\right )dt
\leq  h  \left( \frac{t}\eps+\frac{1}{2}\right)
\| \V^{\eps}(0) \|_{H^1(\mathbb{R})}^2.
\end{equation}
A similar and simpler formula for the next terms is
\begin{equation} \label{eq:sm2}
 \frac{1}{\eps}\int_0^t 
 \sum_{j\in \mathbb{Z}} 
 \left (
\delta_{j}^+(p^{\eps})^2+\delta_{j}^-(p^{\eps}
)^2\right ) \leq   \frac{ h t}\eps 
\|  p^\eps  \|^2_{  H^1(\mathbb R) }
\leq     \frac{ h t}\eps 
\| \V^{\eps}(0) \|_{H^1(\mathbb{R})}^2.
\end{equation}
Next, using the assumption \eqref{geometrie1d} on the mesh and estimate \eqref{estiuepsflux}, one can find a constant $C\geq0$ such that the next term can be  bounded like 
\begin{equation} \label{eq:sm3}
 \int_0^T \underset{j\in \mathbb{Z}}{\sum} \frac{1}{8\eps} \sigma \Delta x_j^2
\frac{(u_{j-\frac12 }^{\eps})^2+(
u_{j+\frac12 }^{\eps})^2}{\eps} \leq C h \|\mathbf{V}^{\eps}(0)\|_{L^2(\mathbb{R})}^2.
\end{equation}
Finally the last term in (\ref{eq:eqn1}) can be bounded using a
Cauchy-Schwarz inequality 
$$
\frac{\sigma^2}{4\eps^3} \sum_{j\in
\mathbb{Z}}\bigg( {\int_{x_{j-\frac12}}^{x_{j+\frac12}}} u^{\eps}(x) dx \bigg)^2
\leq 
\frac{\sigma^2}{4\eps^3}
h \|  \mathbf u^\eps  \|^2_{  L^2(\mathbb R) }
$$
so that
\begin{equation} \label{eq:sm4}
\frac{\sigma^2}{4\eps^3} 
\int_0^t 
 \sum_{j\in
\mathbb{Z}}\bigg( {\int_{x_{j-\frac12}}^{x_{j+\frac12}}} u^{\eps}(x) dx \bigg)^2
\leq 
\frac{\sigma^2}{4\eps^3}
h \|  \mathbf u^\eps  \|^2_{  L^2([0,t]\times \mathbb R) }
\leq 
\frac{\sigma}{4\eps}
h  \|\mathbf{V}^{\eps}(0)\|_{L^2(\mathbb{R})}^2
\end{equation}
by means of the energy identity.
 So using (\ref{eq:sm1}-\ref{eq:sm4}) we obtain for all time $t\leq T$:
\begin{eqnarray*}
\mathscr{E}(t) \leq
\mathscr{E}(0)+h
\left( \frac{t}{\eps}+\frac{1}2 + \frac{t}{\eps}+C+\frac{\sigma
}{4\eps}\right)\| \V^{\eps}(0) \|_{H^1(\mathbb{R})}^2.
\end{eqnarray*}
The initialization stage is consistent so 
 $\mathscr{E}(0)\leq Ch^2\| \V^{\eps}(0) \|_{H^1(\mathbb{R})}^2$.
 Making use of the  initial remark \ref{sigmanotzero}, one   obtains 
\begin{equation*}\label{esti1}
\|\V^{\eps}-\V^{\eps}_h \|_{L^2([0,T]\times \mathbb{R})}
\leq C(T) \sqrt{ \frac{h}{\eps}}
\| \V^{\eps}( 0)
\|_{H^1(\mathbb{R})}, \qquad C(T)\geq0.
\end{equation*}
The proof is ended.
\end{proof}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{proposition}\label{pro:H1}
The bound (\ref{eq:xx}) holds.
\end{proposition}

@@

\begin{proposition}\label{heqtode2d}
The error between the two solutions
can be  upper bounded by
\begin{equation}\label{ine3}
| p^\epsilon-p\|_{   L^\infty([0,T];H^n(\Omega))  }+
\| \uu^\eps 
 \|_{   L^\infty([0,T]; H^n(\Omega))  }
 \leq
C \eps
\|p_0 \|_{H^{3+n}(\Omega)},
\qquad n \in \mathbb N.
\end{equation}
\end{proposition}
\begin{proof}
The structure of the 
proof in the $L^\infty( [0,T];L^2(\Omega))$ norm is the same as the one
of proposition \ref{heqtode}.
Since the coefficients of the problem are constant, similar bounds are obtained
at any order of derivation which proves the estimate for any $n>0$.
\end{proof}

@@

The numerical approximation  
 of the problem  $\mathbf P^\eps$ that we study 
 is the JL-(b) scheme defined
in   \cite{glaceap}
\begin{equation}\label{eq:4}
\mathbf P_h^\eps:
\qquad \left\{ 
\begin{array}{l}
\displaystyle \mid\Omega_{j}\mid\frac{d}{dt} p_{j}^{\eps}+\frac{1}{\varepsilon}
\sum_{r}(l_{jr}\ur^{\eps},\njr)=0\\
\displaystyle \mid
\Omega_{j}\mid\frac{d}{dt}\uj^{\eps}+\frac{1}{\varepsilon}
\sum_{r} l_{jr} \njr    p_{jr}^{\eps}=-\frac{\sigma}{\varepsilon^{2}}\sum_{r}\widehat{
\beta}_{jr}\ur^{\eps},
\end{array}
\right.
\end{equation}
with initial data $p_j^{\eps}(0)=p_0(\mathbf{x}_j)$ and $\mathbf{u}_j^{\eps}(0) = -\eps\sigma^{-1}\nabla p_0(\mathbf{x}_j)$.
The fluxes are defined by the so-called 
corner problem 
\begin{equation}\label{eq:5}
\left\{ \begin{array}{l}
\displaystyle 
p_{jr}^{\eps}-p_j^\eps=
\njr,    \uj^{\eps}-\ur^{\eps}
)-\frac{\sigma}{\varepsilon}( \xr -\xj, \ur^{\eps}), 
\\
\sum_{j}\ljr
p_{jr}^{\eps}\njr=0.
 \end{array}\right.
\end{equation}
This corner problem has been introduced in 
\cite{glaceap} as a multidimensional version
of the 1D Jin-Levermore technique   \cite{jinlev}.
Its solution is provided by the solution
of the  linear system
\begin{equation*} \label{eq:5bis}
\displaystyle \left( \sum_{j}\alj+ 
\sum_{j}\frac{\sigma}{\varepsilon}\bej\right)\ur^{
\eps}= \sum_{j}\ljr
p_j^{\eps}\njr+\sum_j\alj\uj^{\eps},
\end{equation*}
 where the geometry of the mesh is sued to define the  
  matrices  
$\alj$ and $\bej$ 
\begin{equation}\label{eq:6}
\alj=\ljr\njr\otimes\njr,\mbox{ and
}\bej=\ljr\njr\otimes(\xr-\xj) .
\end{equation}
We will use the notations
$A_j=\sum_r\alj$, $A_r=\sum_j\alj$ and
$B_r=\sum_j\bej$. 
By comparison with the scheme $P_h^\eps$ in dimension one, one sees at once
that the multi-dimensional scheme (\ref{eq:4}-\ref{eq:6})  is  more 
tricky than the 1D scheme (\ref{schema1}-\ref{eq:fflux}).

@@

\begin{figure}
\begin{center}
\includegraphics[width=9cm,angle=-90]{sortie_enfin.eps}
\end{center}
\caption{The error is plotted in log scale versus the number of cells per direction for the test problem described
in section \ref{sec::num}.
Each curve corresponds to a value of $\tau\in\{0,\frac14, \frac12, 1, 2  \}$, plus 
a reference line for order one. One sees that the order of convergence is an increasing function
of $\gamma$.}
\label{fig1}
\end{figure}

As predicted by  the theory, the scheme is uniformly AP and the error behavior is a continuous function of $\gamma$
between the  hyperbolic and parabolic limits. However the results are much better, in the sense
the order is greater than the theoretical prediction since
the order is approximatively 1 for $\gamma=0$ (hyperbolic limit)
and 2 for $\gamma=2$ (parabolic regime).
The reason is probably that the theory is based on worst case estimates, as it is often the case for the numerical analysis
of finite volume schemes \cite{FV}.

@@

>>>>

\begin{proposition} \label{heqtode}
There exists a constant $C>0$ such that 
\begin{equation*}
\|\V^{\eps}-\W^\eps\|_{L^\infty([0,T];L^2(\mathbb{R}) )} \leq
C \eps  \|\partial_{xxx}p_0\|_{L^2(\mathbb{R})}.
\end{equation*}
\end{proposition}
\begin{proof}
We define $v=-\frac{\varepsilon}\sigma \partial_x p$ with $p$ the diffusion solution of (\ref{difffff}) and
introduce $R^{\eps}$ such that the solution of the diffusion equation satisfies
\begin{equation}\label{modifdiff}
\left \{
\begin{array}{lll}
\dt p +\frac{1}{\eps}\partial_{x}v=0,\\
\partial_t v+ \frac{1}{\eps}\partial_x p +\frac{\sigma}{\eps^2}v = R^{\eps}
\end{array}
\right.
\end{equation}
where
 $R^{\eps}=\dt v=-\frac{\varepsilon}\sigma \partial_{tx} p=
-\frac{\varepsilon}{\sigma^2} \partial_{xxx}p$. 
Note that 
$\|R^{\eps}(t)\|_{L^2(\mathbb{R})}\leq \|R^{\eps}(0)\|_{L^2(\mathbb{R})}$. 
Denoting $e^{\eps}=p-p^{\eps}$, $f^{\eps}=v-u^{\eps}$, we make the difference between the
systems \eqref{chaleur22} et \eqref{modifdiff}
\begin{equation}\label{diffmodif}
\left \{
\begin{array}{lll}
\dt e^{\eps} +\frac{1}{\eps}\partial_{x}f^{\eps}=0,\\
\partial_t f^{\eps} +\frac{1}{\eps}\partial_x e^{\eps} +\frac{\sigma}{\eps^2}f^{\eps} = R^{\eps}.
\end{array}
\right.
\end{equation}
Since data are well-prepared, one has $e^{\eps}(0)=f^{\eps}(0)=0$.
Consider
$\|\V^{\eps}-\W^\eps\|_{L^2(\mathbb{R})}^2=\|e^{\eps}\|_{L^2(\mathbb{R})}
^2+\|f^{\eps}\|_{L^2(\mathbb{R})}^2$.
Adding the first equation of
\eqref{diffmodif} multiplied by $e^{\eps}$ and the second multiplied by $f^{\eps}$ and integrating
on $\mathbb{R}$,
 we find out that:
$\frac12 \frac{d}{dt}\|\V^{\eps}-\W^\eps\|_{L^2(\mathbb{R})}^2 \leq
\int_{\mathbb{R}}R^{\eps}f^{\eps} dx \leq  \|R^{\eps}\|_{L^2(\mathbb{R})}
\|\V^{\eps}-\W^\eps\|_{L^2(\mathbb{R})}
.$
  The proof is ended by integration between 0 and T.

@@

\|  u^\eps  \|^2_{ L^2( [0,t] : H^1(\mathbb R) }.

Since $\| u^\eps  \|^2_{ L^2( [0,t] : H^1(\mathbb R) }\leq t \| \V^{\eps}(0) \|_{H^1(\mathbb{R})}^2$ and also 
$
\frac{\sigma}
{\eps^2}\| u^\eps  \|^2_{ L^2( [0,t] : H^1(\mathbb R) }\leq  \| \V^{\eps}(0) \|_{H^1(\mathbb{R})}^2$
by (\ref{bee1} ) and (\ref{bee2} ), one gets that
\begin{equation} \label{eq:sm1}
\left( \frac{1}{\eps}+\frac{\sigma}{2\eps^2}\right)
\int_0^t 
\sum_{j\in \mathbb{Z}} 
\left ( 
\delta_{j}^+(u^{\eps})^2+\delta_{j}^-(u^{\eps})^2\right )dt
\leq  h  \left( \frac{t}\eps+\frac{1}{2}\right)
\| \V^{\eps}(0) \|_{H^1(\mathbb{R})}^2.
\end{equation}
A similar and simpler formula for the next terms is
\begin{equation} \label{eq:sm2}
 \frac{1}{\eps}\int_0^t 
 \sum_{j\in \mathbb{Z}} 
 \left (
\delta_{j}^+(p^{\eps})^2+\delta_{j}^-(p^{\eps}
)^2\right ) \leq   \frac{ h t}\eps 
\|  p^\eps  \|^2_{  H^1(\mathbb R) }
\leq     \frac{ h t}\eps 
\| \V^{\eps}(0) \|_{H^1(\mathbb{R})}^2.
\end{equation}
Next, using the assumption \eqref{geometrie1d} on the mesh and estimate \eqref{estiuepsflux}, one can find a constant $C\geq0$ such that the next term can be  bounded like 
\begin{equation} \label{eq:sm3}
 \int_0^T \underset{j\in \mathbb{Z}}{\sum} \frac{1}{8\eps} \sigma \Delta x_j^2
\frac{(u_{j-\frac12 }^{\eps})^2+(
u_{j+\frac12 }^{\eps})^2}{\eps} \leq C h \|\mathbf{V}^{\eps}(0)\|_{L^2(\mathbb{R})}^2.
\end{equation}
Finally the last term in (\ref{eq:eqn1}) can be bounded using a
Cauchy-Schwarz inequality 
$$
\frac{\sigma^2}{4\eps^3} \sum_{j\in
\mathbb{Z}}\bigg( {\int_{x_{j-\frac12}}^{x_{j+\frac12}}} u^{\eps}(x) dx \bigg)^2
\leq 
\frac{\sigma^2}{4\eps^3}
h \|  u^\eps  \|^2_{  L^2(\mathbb R) }
$$
so that
\begin{equation} \label{eq:sm4}
\frac{\sigma^2}{4\eps^3} 
\int_0^t 
 \sum_{j\in
\mathbb{Z}}\bigg( {\int_{x_{j-\frac12}}^{x_{j+\frac12}}} u^{\eps}(x) dx \bigg)^2
\leq 
\frac{\sigma^2}{4\eps^3}
h \| u^\eps  \|^2_{  L^2([0,t]\times \mathbb R) }
\leq 
\frac{\sigma}{4\eps}
h  \|\mathbf{V}^{\eps}(0)\|_{L^2(\mathbb{R})}^2
\end{equation}
by means of the energy identity.
 So using (\ref{eq:sm1}-\ref{eq:sm4}) we obtain for all time $t\leq T$:
\begin{eqnarray*}
\mathscr{E}(t) \leq
\mathscr{E}(0)+h
\left( \frac{t}{\eps}+\frac{1}2 + \frac{t}{\eps}+C+\frac{\sigma
}{4\eps}\right)\| \V^{\eps}(0) \|_{H^1(\mathbb{R})}^2.
\end{eqnarray*}
The initialization stage is consistent so 
 $\mathscr{E}(0)\leq Ch^2\| \V^{\eps}(0) \|_{H^1(\mathbb{R})}^2$.
 Making use of the  initial remark \ref{sigmanotzero}, one   obtains 
\begin{equation*}\label{esti1}
\|\V^{\eps}-\V^{\eps}_h \|_{L^2([0,T]\times \mathbb{R})}
\leq C(T) \sqrt{ \frac{h}{\eps}}
\| \V^{\eps}( 0)
\|_{H^1(\mathbb{R})}, \qquad C(T)\geq0.
\end{equation*}
The proof is ended.
\end{proof}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{lemma}\label{pro:H1}
The bound (\ref{eq:xx}) holds.
\end{lemma}

@@

\begin{proposition}\label{heqtode2d}
The error between the two solutions
can be  upper bounded by
\begin{equation}\label{ine3}
\| p^\epsilon-p\|_{   L^\infty([0,T];H^n(\Omega))  }+
\| \mathbf{v} 
 \|_{   L^\infty([0,T]; H^n(\Omega))  }
 \leq
C \eps
\|p_0 \|_{H^{3+n}(\Omega)},
\qquad n \in \mathbb N.
\end{equation}
\end{proposition}
\begin{proof}
The structure of the 
proof in the $L^\infty( [0,T];L^2(\Omega))$ norm is the same as the one
of proposition \ref{heqtode}.
Since the coefficients of the problem are constant, similar bounds are obtained
at any order of derivation which proves the estimate for any $n>0$.
\end{proof}

@@

The numerical approximation  
 of the problem  $\mathbf P^\eps$ that we study 
 is the JL-(b) scheme defined
in   \cite{glaceap}
\begin{equation}\label{eq:4}
\mathbf P_h^\eps:
\qquad \left\{ 
\begin{array}{l}
\displaystyle \mid\Omega_{j}\mid\frac{d}{dt} p_{j}^{\eps}+\frac{1}{\varepsilon}
\sum_{r}(l_{jr}\ur^{\eps},\njr)=0\\
\displaystyle \mid
\Omega_{j}\mid\frac{d}{dt}\uj^{\eps}+\frac{1}{\varepsilon}
\sum_{r} l_{jr} \njr    p_{jr}^{\eps}=-\frac{\sigma}{\varepsilon^{2}}\sum_{r}\widehat{
\beta}_{jr}\ur^{\eps},
\end{array}
\right.
\end{equation}
with initial data $p_j^{\eps}(0)=p_0(\mathbf{x}_j)$ and $\mathbf{u}_j^{\eps}(0) = -\eps\sigma^{-1}\nabla p_0(\mathbf{x}_j)$.
The fluxes are defined by the so-called 
corner problem 
\begin{equation}\label{eq:5}
\left\{ \begin{array}{l}
\displaystyle 
p_{jr}^{\eps}-p_j^\eps=
(\njr,    \uj^{\eps}-\ur^{\eps}
)-\frac{\sigma}{\varepsilon}( \xr -\xj, \ur^{\eps}), 
\\
\sum_{j}\ljr
p_{jr}^{\eps}\njr=0.
 \end{array}\right.
\end{equation}
This corner problem has been introduced in 
\cite{glaceap} as a multidimensional version
of the 1D Jin-Levermore technique   \cite{jinlev}.
Its solution is provided by the solution
of the  linear system
\begin{equation*} \label{eq:5bis}
\displaystyle \left( \sum_{j}\alj+ 
\sum_{j}\frac{\sigma}{\varepsilon}\bej\right)\ur^{
\eps}= \sum_{j}\ljr
p_j^{\eps}\njr+\sum_j\alj\uj^{\eps},
\end{equation*}
 where the geometry of the mesh is sued to define the  
  matrices  
$\alj$ and $\bej$ 
\begin{equation}\label{eq:6}
\alj=\ljr\njr\otimes\njr,\mbox{ and
}\bej=\ljr\njr\otimes(\xr-\xj) .
\end{equation}
We will use the notations
$A_j=\sum_r\alj$, $A_r=\sum_j\alj$ and
$B_r=\sum_j\bej$. 
By comparison with the scheme $P_h^\eps$ in dimension one, one sees at once
that the multi-dimensional scheme (\ref{eq:4}-\ref{eq:6})  is  more 
tricky than the 1D scheme (\ref{schema1}-\ref{eq:fflux}).

@@

\begin{figure}
\begin{center}
\includegraphics[width=9cm,angle=-90]{sortie_enfin.eps}
\end{center}
\caption{The error is plotted in log scale versus the number of cells per direction for the test problem described
in section \ref{sec::num}.
Each curve corresponds to a value of $\tau\in\{0,\frac14, \frac12, 1, 2  \}$, plus 
a reference line for order one. One sees that the order of convergence is an increasing function
of $\tau$.}
\label{fig1}
\end{figure}

As predicted by  the theory, the scheme is uniformly AP and the error behavior is a continuous function of $\gamma$
between the  hyperbolic and parabolic limits. However the results are much better, in the sense
the order is greater than the theoretical prediction since
the order is approximatively 1 for $\gamma=0$ (hyperbolic limit)
and 2 for $\gamma=2$ (parabolic regime).  We can find a closed result on the second order convergence for the parabolic regime in the paper \cite{TAHO} (1D linear problem).
The reason is probably that the theory is based on worst case estimates, as it is often the case for the numerical analysis
of finite volume schemes \cite{FV}.

@@

\bibitem{TAHO} D. Aregba, M. Briani, R. Natalini, Time Asymptotic High Order Schemes for Dissipative BGK Hyperbolic Systems, arXiv:1207.6279.
